---
title: "ML - German_Credit_Risk"
author: "Bryzik Michał, Gruca Kacper"
date: "2024-01-09"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
  html_notebook:
    toc: true
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

## Random Forest - one by one

Let's start with cleaning the environment and loading the data.

```{r}
rm(list=ls())

df <- read.csv("./data/output/df.csv")
df.train <- read.csv("./data/output/df.train.csv")
df.test <- read.csv("./data/output/df.test.csv")

df.train$class <- factor(df.train$class, levels = c(0, 1), labels = c("bad", "good"))
df.test$class <- factor(df.test$class, levels = c(0, 1), labels = c("bad", "good"))
```

Prepare df as before.

```{r}
rfe_columns <- read.csv("./data/output/rfe_columns.csv")
rfe_columns <- c("class", rfe_columns$x)

df.train <- df.train[, rfe_columns]
df.test <- df.test[, rfe_columns]
```

Estimate first - basic model with all variables and default parameters

```{r, message=FALSE}
library(randomForest)
library(caret)
library(pROC)
library(dplyr)
```


```{r}
set.seed(123456789)
rf <- randomForest(class ~ ., 
                   data = df.train)

saveRDS(rf, file = "./data/output/rf/rf.RDS")
```

```{r}
print(rf)
```

Estimated out-of-bag-error: 13.57%

```{r}
plot(rf)
```

The plot above helps determine the appropriate number of trees.

The dark solid line presents the total OOB error, the red line presents the prediction error for "good" response, while the green line for "bad" response.

As the number of trees is higher, the OOB error is smaller, and converges to approx. 120 trees.

Hence, let us try to limit number of trees and estimate the model on bootstrap samples from the full data set.

We also increase the number of predictors used from 5 to 10.

```{r}
set.seed(123456789)
rf2 <- 
  randomForest(class ~ .,
               data = df.train,
               ntree = 120,
               sampsize = nrow(df.train),
               mtry = 10,
               # minimum number of obs in the terminal nodes
               nodesize = 100,
               # we also generate predictors importance measures,
               importance = TRUE)

saveRDS(rf2, file = "./data/output/rf/rf2.RDS")
print(rf2)
```

Let us try to optimize the mtry parameter using the cross-validation process.

We consider 35 predictors in the model formula. Hence, let us try mtry between 2 and 31.

```{r, max.height='150px'}
parameters_rf <- expand.grid(mtry = 2:31)
ctrl_oob <- trainControl(method = "oob", classProbs = TRUE)

set.seed(123456789)

rf3 <-
  train(class ~ .,
        data = df.train,
        method = "rf",
        ntree = 120,
        nodesize = 100,
        tuneGrid = parameters_rf,
        trControl = ctrl_oob,
        importance = TRUE)

rf3
saveRDS(rf3, file = "./data/output/rf/rf3.RDS") 
```

```{r}
plot(rf3$results$mtry,
     rf3$results$Accuracy, type = "b")
```

```{r}
plot(rf3$results$mtry,
     rf3$results$Kappa, type = "b")
```

```{r}
modelLookup("ranger")
```

Let's apply grid search and try to optimize mtry and min.node.size parameters

```{r}
parameters_ranger <- 
  expand.grid(mtry = 4:15,
              # split rule
              splitrule = "gini",
              # minimum size of the terminal node
              min.node.size = c(50, 100, 150))

ctrl_cv5 <- trainControl(method = "cv", 
                         number =    5,
                         classProbs = T)

set.seed(123456789)
rf3a <- 
  train(class ~ ., 
        data = df.train, 
        method = "ranger", 
        num.trees = 120, # default = 500
        # numbers of processor cores to use in computations
        num.threads = 3,
        # impurity measure
        importance = "impurity",
        # parameters
        tuneGrid = parameters_ranger, 
        trControl = ctrl_cv5)

saveRDS(rf3a, file = "./data/output/rf/rf3a.RDS") 
```

```{r, max.height='150px'}
rf3a
```

```{r}
plot(rf3a)
```

In this case the omptimal value for mtry is 6.

Let's compare all of the models.

```{r, message=FALSE}
pred.train.rf <- predict(rf, 
                         df.train, 
                         type = "prob")[, "good"]
ROC.train.rf  <- roc(as.numeric(df.train$class == "good"), 
                     pred.train.rf)


pred.test.rf  <- predict(rf, 
                         df.test, 
                         type = "prob")[, "good"]
ROC.test.rf   <- roc(as.numeric(df.test$class == "good"), 
                     pred.test.rf)


pred.train.rf2 <- predict(rf2, 
                          df.train, 
                          type = "prob")[, "good"]
ROC.train.rf2  <- roc(as.numeric(df.train$class == "good"), 
                      pred.train.rf2)


pred.test.rf2  <- predict(rf2, 
                          df.test, 
                          type = "prob")[, "good"]
ROC.test.rf2   <- roc(as.numeric(df.test$class == "good"), 
                      pred.test.rf2)


pred.train.rf3 <- predict(rf3, 
                          df.train, 
                          type = "prob")[, "good"]
ROC.train.rf3  <- roc(as.numeric(df.train$class == "good"), 
                      pred.train.rf3)


pred.test.rf3  <- predict(rf3, 
                          df.test, 
                          type = "prob")[, "good"]
ROC.test.rf3   <- roc(as.numeric(df.test$class == "good"), 
                      pred.test.rf3)


pred.train.rf3a <- predict(rf3a, 
                           df.train, 
                           type = "prob")[, "good"]
ROC.train.rf3a  <- roc(as.numeric(df.train$class == "good"), 
                       pred.train.rf3a)


pred.test.rf3a  <- predict(rf3a, 
                           df.test, 
                           type = "prob")[, "good"]
ROC.test.rf3a   <- roc(as.numeric(df.test$class == "good"), 
                       pred.test.rf3a)
```

```{r}
list(
  ROC.train.rf   = ROC.train.rf,
  ROC.test.rf    = ROC.test.rf,
  ROC.train.rf2  = ROC.train.rf2,
  ROC.test.rf2   = ROC.test.rf2,
  ROC.train.rf3  = ROC.train.rf3,
  ROC.test.rf3   = ROC.test.rf3,
  ROC.train.rf3a = ROC.train.rf3a,
  ROC.test.rf3a  = ROC.test.rf3a
) %>%
  pROC::ggroc(alpha = 0.5, linetype = 1, size = 1) + 
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), 
               color = "grey", 
               linetype = "dashed") +
  labs(title = paste0("Gini TEST: ",
                      "rf = ", 
                      round(100 * (2 * auc(ROC.test.rf) - 1), 1), "%, ",
                      "rf2 = ", 
                      round(100 * (2 * auc(ROC.test.rf2) - 1), 1), "%, ",
                      "rf3 = ", 
                      round(100 * (2 * auc(ROC.test.rf3) - 1), 1), "%, ",
                      "rf3a = ", 
                      round(100 * (2 * auc(ROC.test.rf3a) - 1), 1), "% "),
       subtitle =  paste0("Gini TRAIN: ",
                          "rf = ", 
                          round(100 * (2 * auc(ROC.train.rf) - 1), 1), "%, ",
                          "rf2 = ", 
                          round(100 * (2 * auc(ROC.train.rf2) - 1), 1), "%, ",
                          "rf3 = ", 
                          round(100 * (2 * auc(ROC.train.rf3) - 1), 1), "%, ",
                          "rf3a = ", 
                          round(100 * (2 * auc(ROC.train.rf3a) - 1), 1), "% ")) +
  theme_bw() + coord_fixed() +
  scale_color_brewer(palette = "Paired")
```

Finally the random forest model with grid search (rf3a) with mtry = 6, splitrule = gini and min.node.size = 50 turned out to be the best.




