---
title: "ML - German_Credit_Risk"
author: "Bryzik Michał, Gruca Kacper"
date: "2024-01-09"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
  html_notebook:
    toc: true
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

## GMB

In this section we will perform GBM model.
Let's start with clearing the environment and preparing df.

```{r, message=FALSE}
library(gbm)
library(dplyr)
library(caret)
```


```{r}
rm(list=ls())

df <- read.csv("./data/output/df.csv")
df.train <- read.csv("./data/output/df.train.csv")
df.test <- read.csv("./data/output/df.test.csv")
```

```{r}
rfe_columns <- read.csv("./data/output/rfe_columns.csv")
rfe_columns <- c("class", rfe_columns$x)

df.train <- df.train[, rfe_columns]
df.test <- df.test[, rfe_columns]
```

```{r}
# Make factor
df$class <- factor(df$class, levels = c(0, 1), labels = c("bad", "good"))
df.train$class <- factor(df.train$class, levels = c(0, 1), labels = c("bad", "good"))
df.test$class <- factor(df.test$class, levels = c(0, 1), labels = c("bad", "good"))
```

Function to define model formula

```{r}
# define model formula - all variables apart from class
create_model_formula <- function(df, explained_variable){
  
  model_formula <- paste(explained_variable, "~")
  
  for (i in 2:length(colnames(df))){
    if(i == 2){
      model_formula <- paste(model_formula, colnames(df)[i])
    } else {
      model_formula <- paste(model_formula, " + ", colnames(df)[i])
    }
  }
  
  return(as.formula(model_formula))
}
```

Define 2 model formulas - one for class (factor) and one for class_1 (number).

```{r}
model_formula01 <- create_model_formula(df.train, "class_1")
model_formula <- create_model_formula(df.train, "class")
```

GBM requires target variable to be defined as 0 and 1, not factor so we do not change it. However we will also need variable "class" with factor.
Define new variable - class_1 with 0 and 1.

```{r}
df.train$class_1 <- (df.train$class == "good") * 1
df.test$class_1 <- (df.test$class == "good") * 1
```


### Simple GBM

In this section we will build simple GBM modelm without any hyperparameter tuning.

```{r}
set.seed(123456789)
gbm.1 <- 
  gbm(model_formula01,
      data = df.train,
      distribution = "bernoulli",
      # total number of trees
      n.trees = 500,
      # number of variable interactions - actually depth of the trees
      interaction.depth = 4,
      # shrinkage parameter - speed (pace) of learning
      shrinkage = 0.01,
      verbose = FALSE)

gbm.1 %>% saveRDS("./data/output/gbm/gbm.1.rds")

# predict train
df.pred.train.gbm <- predict(gbm.1,
                                  df.train, 
                                  type = "response",
                                  n.trees = 500)
# predict test
df.pred.test.gbm <- predict(gbm.1,
                             df.test, 
                             type = "response",
                             n.trees = 500)
```

```{r, message=FALSE, warning=FALSE}
# Load function to get accuracy and gini values
source("./scripts/getAccuracyAndGini2.R")

# Training set
getAccuracyAndGini2(data = data.frame(class = df.train$class,
                                      pred = df.pred.train.gbm),
                    predicted_probs = "pred",
                    target_variable = "class",
                    target_levels = c("good", "bad"),
                    predicted_class = "good")
```

```{r, warning=FALSE}
# Test set
getAccuracyAndGini2(data = data.frame(class = df.test$class,
                                      pred = df.pred.test.gbm),
                    predicted_probs = "pred",
                    target_variable = "class",
                    target_levels = c("good", "bad"),
                    predicted_class = "good")
```


```{r, message=FALSE}
ROC.train.gbm <- pROC::roc(df.train$class, 
                           df.pred.train.gbm)

ROC.test.gbm  <- pROC::roc(df.test$class, 
                           df.pred.test.gbm)

cat("AUC for train = ", pROC::auc(ROC.train.gbm), 
    ", Gini for train = ", 2 * pROC::auc(ROC.train.gbm) - 1, "\n", sep = "")

cat("AUC for test = ", pROC::auc(ROC.test.gbm), 
    ", Gini for test = ", 2 * pROC::auc(ROC.test.gbm) - 1, "\n", sep = "")
```

Plot ROC

```{r}
list(
  ROC.train.gbm = ROC.train.gbm,
  ROC.test.gbm  = ROC.test.gbm
) %>%
  pROC::ggroc(alpha = 0.5, linetype = 1, size = 1) + 
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), 
               color = "grey", 
               linetype = "dashed") +
  labs(subtitle = paste0("Gini TRAIN: ",
                         "gbm = ", 
                         round(100 * (2 * auc(ROC.train.gbm) - 1), 1), "%, ",
                         "Gini TEST: ",
                         "gbm = ", 
                         round(100 * (2 * auc(ROC.test.gbm) - 1), 1), "%, "
  )) +
  theme_bw() + coord_fixed() +
  scale_color_brewer(palette = "Paired")
```

### Hyperparameter tuning

```{r}
parameters_gbm <- expand.grid(interaction.depth = c(1, 2, 4),
                              n.trees = c(100, 300, 500),
                              shrinkage = c(0.01, 0.1), 
                              n.minobsinnode = c(150, 250, 400))

ctrl_cv3 <- trainControl(method = "cv", 
                         number = 10,
                         classProbs = TRUE,
                         summaryFunction = twoClassSummary)
```

```{r, max.height='150px', warning=FALSE}
set.seed(123456789)
gbm.2  <- train(model_formula,
                       data = df.train,
                       distribution = "bernoulli",
                       method = "gbm",
                       tuneGrid = parameters_gbm,
                       trControl = ctrl_cv3,
                       verbose = FALSE)

gbm.2
```

```{r}
res <- gbm.2$results

saveRDS(object = gbm.2,
        file   = "./data/output/gbm/gbm.2.rds")
```

```{r}
# Predict train
df.pred.train.gbm2 <- predict(gbm.2,
                             df.train, 
                             type = "prob",
                             n.trees = 500)
# Predict test
df.pred.test.gbm2 <- predict(gbm.2,
                            df.test, 
                            type = "prob",
                            n.trees = 500)
```

```{r, warning=FALSE}
# Training set
accuracy_gini_train <- getAccuracyAndGini2(data = data.frame(class = df.train$class,
                                      pred = df.pred.train.gbm2[, "good"]),
                    predicted_probs = "pred",
                    target_variable = "class",
                    target_levels = c("good", "bad"),
                    predicted_class = "good")
```

```{r, warning=FALSE}
# Test set
accuracy_gini_test <- getAccuracyAndGini2(data = data.frame(class = df.test$class,
                                      pred = df.pred.test.gbm2[, "good"]),
                    predicted_probs = "pred",
                    target_variable = "class",
                    target_levels = c("good", "bad"),
                    predicted_class = "good")
```

```{r}
write.csv(accuracy_gini_test, "./data/accuracy_gini_test.csv")
write.csv(accuracy_gini_train, "./data/accuracy_gini_train.csv")
```

```{r, message=FALSE}
ROC.train.gbm2 <- pROC::roc(df.train$class, 
                           df.pred.train.gbm2[, "good"])

ROC.test.gbm2  <- pROC::roc(df.test$class, 
                           df.pred.test.gbm2[, "good"])
```

```{r}
cat("AUC for train = ", pROC::auc(ROC.train.gbm2), 
    ", Gini for train = ", 2 * pROC::auc(ROC.train.gbm2) - 1, "\n", sep = "")

cat("AUC for test = ", pROC::auc(ROC.test.gbm2), 
    ", Gini for test = ", 2 * pROC::auc(ROC.test.gbm2) - 1, "\n", sep = "")
```

```{r}
list(
  ROC.train.gbm2 = ROC.train.gbm2,
  ROC.test.gbm2  = ROC.test.gbm2
) %>%
  pROC::ggroc(alpha = 0.5, linetype = 1, size = 1) + 
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), 
               color = "grey", 
               linetype = "dashed") +
  labs(subtitle = paste0("Gini TRAIN: ",
                         "gbm = ", 
                         round(100 * (2 * auc(ROC.train.gbm2) - 1), 1), "%, ",
                         "Gini TEST: ",
                         "gbm = ", 
                         round(100 * (2 * auc(ROC.test.gbm2) - 1), 1), "%, "
  )) +
  theme_bw() + coord_fixed() +
  scale_color_brewer(palette = "Paired")
```













